---
title: "Board Game Geek Analysis"
author: "Bill Best"
date: "February 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RSQLite)
library(ggplot2)
library(dplyr)
library(tidyr)
library(corrplot)
library(stats)
```


```{r load, echo=FALSE}
get.data.from.sqlite <- function(db.name, table.name) {
  con <- dbConnect(SQLite(), dbname=db.name)
  db.query <- paste('select * from', table.name, ";")
  table.data <- dbGetQuery(con, db.query)
  return(table.data)
}

game.data <- get.data.from.sqlite("bgg_data_3.db", "board_games")

# Then clean
# there is a minimum set of data I want to work with

clean.data <- game.data %>% 
  filter(grepl('[[:digit:]]+', rank)) %>% mutate(rank = as.numeric((rank))) %>%
  filter(grepl('[[:digit:]]+', num_votes)) %>% mutate(num_votes = as.numeric((num_votes))) %>%
  filter(grepl('[[:digit:].]+', geek_rating)) %>% mutate(geek_rating = as.numeric((geek_rating))) %>%
  filter(grepl('[[:digit:].]+', avg_rating)) %>% mutate(avg_rating = as.numeric((avg_rating))) %>%
  filter(grepl('[[:digit:].]+', avg_rating_std_deviations)) %>% mutate(avg_rating_std_deviations = as.numeric((avg_rating_std_deviations)))

```

## Plot Votes vs Comments with Rank Page

As rank increases, people vote more, and comment more. People are more interested in talking about things they like.
Note that this is log-log scale.

```{r comment.vote.rank, echo=FALSE}

valid.data <- clean.data %>% filter(grepl("[[:digit:]]+", num_comments)) %>% mutate(num_comments = as.numeric(num_comments))

votes.comments <- valid.data %>% select(num_votes, num_comments, rank)
votes.comments <- votes.comments[complete.cases(votes.comments), ]

ggplot(votes.comments, aes(x=num_votes, y=num_comments, color=rank)) + geom_point() + 
  scale_colour_continuous(limits= c(1, 5000), low="green", high="blue") + 
  scale_x_log10() + scale_y_log10() + xlab("# Votes") + ylab("# Comments")
```

## How correlated are geek_rating and rank?

Very strongly. abs(Corr) = 0.88

```{r ,echo=FALSE}

cor(select(clean.data, c(rank, geek_rating)))

```

## Are Number of Fans a good indicator of geek rating?

Actually, no. Number of fans is not closely correlated.

```{r num.fans.check, echo=FALSE}

fan.data <- clean.data %>% 
  filter(grepl("[[:digit:]]+", num_fans)) %>% mutate(num_fans = as.numeric(num_fans)) %>% filter(num_fans >  0) %>%
  filter(grepl("[[:digit:]]+", num_comments)) %>% mutate(num_comments = as.numeric(num_comments)) %>% filter(num_comments >  0) %>%
  filter(grepl("[[:digit:].]+", weight)) %>% mutate(weight = as.numeric(weight)) %>% filter(weight >  0) %>%
  mutate(rank = 5000-rank)

fan.data <- fan.data %>% 
  select(Rank = rank, "Geek Rating" = geek_rating, "Avg. Rating" = avg_rating, "# Fans" = num_fans, "# Votes" = num_votes, "# Comments" = num_comments, "Weight" = weight)

f.cor <- cor(fan.data)
f.cor
corrplot(f.cor)


fan.data <- clean.data %>% 
  filter(grepl("[[:digit:]]+", num_fans)) %>% mutate(num_fans = as.numeric(num_fans)) %>% filter(num_fans >  0) %>%
  filter(grepl("[[:digit:]]+", num_comments)) %>% mutate(num_comments = as.numeric(num_comments)) %>% filter(num_comments >  0) %>%
  filter(grepl("[[:digit:].]+", weight)) %>% mutate(weight = as.numeric(weight)) %>% filter(weight >  0) %>%
  mutate(rank = 5000-rank)

fan.data <- fan.data %>% 
  mutate(num_fans = log10(num_fans), num_votes = log10(num_votes), num_comments = log10(num_comments)) %>%
  select(Rank = rank, "Geek Rating" = geek_rating, "Avg. Rating" = avg_rating, "log10(# Fans)" = num_fans, "log10(# Votes)" = num_votes, "log10(# Comments)" = num_comments, "Weight" = weight)


f.cor.loglog <- cor(fan.data)
f.cor.loglog
corrplot(f.cor.loglog)


fan.data <- clean.data %>%
  filter(grepl("[[:digit:]]+", num_fans)) %>% mutate(num_fans = as.numeric(num_fans)) %>% filter(num_fans >  0) %>%
  filter(grepl("[[:digit:]]+", num_expansions)) %>% mutate(num_expansions = as.numeric(num_expansions)) %>%
  filter(grepl("[[:digit:]]+", owned)) %>% mutate(owned = as.numeric(owned)) %>%
  filter(grepl("[[:digit:]]+", prev_owned)) %>% mutate(prev_owned = as.numeric(prev_owned)) %>%
  mutate(rank = 5000-rank)

fan.data <- fan.data %>% 
  mutate(`Est. Sales` = owned + prev_owned) %>%
  select(Rank = rank, "Geek Rating" = geek_rating, "Avg. Rating" = avg_rating, "# Expansions" = num_expansions, num_fans)

f.cor <- cor(fan.data)

corrplot(f.cor)


fan.data <- clean.data %>% 
  filter(grepl("[[:digit:]]+", num_fans)) %>% mutate(num_fans = as.numeric(num_fans)) %>% filter(num_fans >  0) %>%
  filter(grepl("[[:digit:]]+", num_comments)) %>% mutate(num_comments = as.numeric(num_comments)) %>% filter(num_comments >  0) %>%
  filter(grepl("[[:digit:].]+", weight)) %>% mutate(weight = as.numeric(weight)) %>% filter(weight >  0) %>%
  mutate(rank = 5000-rank)

fan.data <- fan.data %>% 
  mutate(`FV Score` = (log10(num_votes) + avg_rating )) %>%
  mutate(num_votes = log10(num_votes), avg_rating = avg_rating, `FV Score` = `FV Score`) %>%
  select(Rank = rank, "log10(# Votes)" = num_votes, avg_rating, `FV Score`)


f.cor.cv.score <- cor(fan.data)
f.cor.cv.score




```


```{r, echo=FALSE}

fan.data <- clean.data %>% 
  filter(grepl("[[:digit:]]+", num_fans)) %>% mutate(num_fans = as.numeric(num_fans)) %>% filter(num_fans >  0) %>%
  filter(grepl("[[:digit:]]+", num_comments)) %>% mutate(num_comments = as.numeric(num_comments)) %>% filter(num_comments >  0) %>%
  mutate(rank = 5000-rank)

fan.data <- fan.data %>% 
  mutate(num_fans = log10(num_fans), num_votes = log10(num_votes), num_comments = log10(num_comments)) %>%
  select(Rank = rank, "Geek Rating" = geek_rating, "Avg. Rating" = avg_rating, "log10(# Fans)" = num_fans, "log10(# Votes)" = num_votes, "log10(# Comments)" = num_comments)

```


## Weight Analysis

People generally prefer more complex games over simpler games,

```{r weight.avg.Rating, echo=FALSE}

valid.data <- clean.data %>% filter(grepl("[[:digit:].]+", weight)) %>% mutate(weight = as.numeric(weight))

weight.rating <- valid.data %>% select(weight, avg_rating, rank)
weight.rating <- weight.rating[complete.cases(weight.rating), ]

ggplot(weight.rating, aes(x=weight, y=avg_rating, color=rank)) + geom_point() + scale_colour_continuous("Rank", limits= c(1, 5000), low="green", high="blue")


```

## Non-Rentention Rates

As Geek rating increases, people are more likely to keep the board games they own. That is, the "non-retention" rate, rate of previous ownership, goes down.


```{r retention, echo=FALSE}

valid.data <- clean.data %>% 
  filter(grepl("[[:digit:]]+", owned)) %>% mutate(owned = as.numeric(owned)) %>%
  filter(grepl("[[:digit:]]+", prev_owned)) %>% mutate(prev_owned = as.numeric(prev_owned))
  

retention.rank <- valid.data %>% mutate(retention = prev_owned / (owned + prev_owned)) %>% select(retention, geek_rating, rank)
retention.rank <- retention.rank[complete.cases(retention.rank), ]

ggplot(retention.rank, aes(x=geek_rating, y=retention, color=rank)) + geom_point() + scale_colour_continuous("Rank", limits= c(1, 5000), low="green", high="blue") + xlab("Geek Rating") + ylab("Non-Retention")

```


## Find Top Mechanics and Categories

I used log10(fans) for each game, and took the mean of that to evaluate and rank the categorical variables.


```{r top.mechanics, echo=FALSE}

valid.data <- clean.data %>% filter(grepl("[[:digit:].]+", num_fans)) %>% mutate(fans = as.numeric(num_fans))

fans.mechanics <- valid.data %>% select(fans, num_votes, mechanics) %>% 
  select(fans, mechanics)

fans.mechanics <- fans.mechanics %>% 
  mutate(mechanics = strsplit(mechanics, split = "|", fixed=TRUE)) %>% unnest(mechanics) %>%
  mutate(fans = log10(fans)) %>%
  group_by(mechanics) %>% summarise(fans = mean(fans)) %>% arrange(desc(fans))

top.mechanics <- fans.mechanics[1:10, ]
top.mechanics[, 1]

```

```{r top.categories, echo=FALSE}

valid.data <- clean.data %>% filter(grepl("[[:digit:].]+", num_fans)) %>% mutate(fans = as.numeric(num_fans))

# how far above average geek rating is the game? <- this gives us a score
# how many people thought talking about it was important? <- this gives us weight
fans.categories <- valid.data %>% select(fans, num_votes, categories) %>% 
  select(fans, categories)

fans.categories <- fans.categories %>% 
  mutate(categories = strsplit(categories, split = "|", fixed=TRUE)) %>% unnest(categories) %>%
  mutate(fans = log10(fans)) %>%
  group_by(categories) %>% summarise(fans = mean(fans)) %>% arrange(desc(fans))

top.categories <- fans.categories[1:10, ]
top.categories[, 1]

```

```{r m.c.counts, echo=FALSE}

# Number of top mechanics/categories per game
count.top.mechanics <- 
  clean.data %>%
  mutate(mechanics = strsplit(mechanics, split = "|", fixed=TRUE)) %>% unnest(mechanics) %>%
  filter(mechanics %in% top.mechanics$mechanics) %>%
  group_by(game_id) %>% summarise(num.top.mechanics = n())

count.top.categories <- 
  clean.data %>%
  mutate(categories = strsplit(categories, split = "|", fixed=TRUE)) %>% unnest(categories) %>%
  group_by(game_id) %>% summarise(num.top.categories = n())

# Total mechanics/categories per game
count.mechanics <- 
  clean.data %>%
  mutate(mechanics = strsplit(mechanics, split = "|", fixed=TRUE)) %>% unnest(mechanics) %>%
  filter(mechanics %in% top.mechanics$mechanics) %>%
  group_by(game_id) %>% summarise(num.mechanics = n())

count.categories <- 
  clean.data %>%
  mutate(categories = strsplit(categories, split = "|", fixed=TRUE)) %>% unnest(categories) %>%
  group_by(game_id) %>% summarise(num.categories = n())

```

```{r ,echo=FALSE}

mechanic.data <- clean.data %>% mutate(mechanics = strsplit(mechanics, split = "|", fixed=TRUE)) %>% unnest(mechanics) %>%
  select(mechanics, geek_rating, avg_rating, num_votes, rank)
category.data <- clean.data %>% mutate(categories = strsplit(categories, split = "|", fixed=TRUE)) %>% unnest(categories) %>%
  select(categories, geek_rating, avg_rating, num_votes, rank)

mech.mod <- lm(geek_rating ~ mechanics, mechanic.data)

anova(mech.mod)

```

```{r top.mechanics, echo=FALSE}

valid.data <- clean.data %>% filter(grepl("[[:digit:].]+", num_fans)) %>% mutate(fans = as.numeric(num_fans))

fans.mechanics <- valid.data %>% select(num_votes, avg_rating, mechanics)

fans.mechanics <- fans.mechanics %>% 
  mutate(mechanics = strsplit(mechanics, split = "|", fixed=TRUE)) %>% unnest(mechanics) %>%
  mutate(score = (log10(num_votes) + avg_rating )) %>%
  group_by(mechanics) %>% summarise(score = mean(score)) %>% arrange((score))

top.mechanics <- fans.mechanics[1:10, ]
top.mechanics[, 1]

```


```{r top.categories, echo=FALSE}

valid.data <- clean.data %>% filter(grepl("[[:digit:].]+", num_fans)) %>% mutate(fans = as.numeric(num_fans))

fans.categories <- valid.data %>% select(num_votes, avg_rating, categories)

fans.categories <- fans.categories %>% 
  mutate(categories = strsplit(categories, split = "|", fixed=TRUE)) %>% unnest(categories) %>%
  mutate(score = (log10(num_votes) + avg_rating )) %>%
  group_by(categories) %>% summarise(score = mean(score)) %>% arrange((score))

fans.categories <- fans.categories[1:10, ]
fans.categories[, 1]

```

```{r fan.categories, echo=FALSE}

valid.data <- clean.data %>% filter(grepl("[[:digit:].]+", num_fans)) %>% mutate(fans = as.numeric(num_fans)) %>% filter(fans > 0)

fans.categories <- valid.data %>% select(num_votes, fans, categories)

fans.categories <- fans.categories %>% 
  mutate(categories = strsplit(categories, split = "|", fixed=TRUE)) %>% unnest(categories) %>%
  mutate(fans = log10(fans)) %>%
  group_by(categories) %>% summarise(fans = mean(fans)) %>% arrange(desc(fans))

fans.categories <- fans.categories[1:10, ]
fans.categories[, 1]

```

```{r fan.mechs, echo=FALSE}

valid.data <- clean.data %>% filter(grepl("[[:digit:].]+", num_fans)) %>% mutate(fans = as.numeric(num_fans)) %>% filter(fans > 0)

fans.mechanics <- valid.data %>% select(num_votes, fans, mechanics)

fans.mechanics <- fans.mechanics %>% 
  mutate(mechanics = strsplit(mechanics, split = "|", fixed=TRUE)) %>% unnest(mechanics) %>%
  mutate(fans = log10(fans)) %>%
  group_by(mechanics) %>% summarise(fans = mean(fans)) %>% arrange((fans))

fans.mechanics <- fans.mechanics[1:10, ]
fans.mechanics[, 1]

```
